MINOS Handover Plots

### 2024-07-11 minos/validation/handovers.Rmd
Added reflection for SF_12 model and some other small things

```{r}
# Read raw datafiles in
#raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.files <- list.files(here::here('data', 'imputed_final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
#raw.files <- list.files(here::here('data', 'scaled_scotland_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))
raw.dat <- raw.dat %>%
  filter(weight > 0)

out.path <- here::here('output', 'default_config/')
#out.path <- here::here('output', 'scotland_scaled/')
base.dat <- read_singular_local_out(out.path, 'baseline', 
                                    drop.dead = TRUE, 
                                    drop.zero.weight = TRUE)
#base.dat <- read_singular_local_out(out.path, 'goodHeatingDummy', drop.dead = TRUE)
```

@@

```{r}
# Step 1: Filter for households with children
households_with_kids <- raw.dat %>%
  filter(nkids > 0) %>%
  select(hidp, time, hh_income) %>%
  distinct() %>%
  select(-hidp)

# Step 2: Calculate income quintiles
households_with_kids <- households_with_kids %>%
  group_by(time) %>%
  mutate(income_quintile = ntile(hh_income, 5)) %>%
  ungroup()

# Step 3: Count households within each income quintile per year
household_count_by_income_quintile <- households_with_kids %>%
  group_by(time, income_quintile) %>%
  summarise(n = n())

household_count_by_income_quintile
```

>>>>

```{r}
# Read raw datafiles in
#raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
#raw.files <- list.files(here::here('data', 'imputed_final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.files <- list.files(here::here('data', 'scaled_scotland_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))
raw.dat <- raw.dat %>%
  filter(weight > 0)

#out.path <- here::here('output', 'default_config/')
out.path <- here::here('output', 'scotland_scaled/')
base.dat <- read_singular_local_out(out.path, 'baseline', 
                                    drop.dead = TRUE, 
                                    drop.zero.weight = TRUE)
#base.dat <- read_singular_local_out(out.path, 'goodHeatingDummy', drop.dead = TRUE)
```

@@

```{r}
# # Step 1: Filter for households with children
# households_with_kids <- raw.dat %>%
#   filter(nkids > 0) %>%
#   select(hidp, time, hh_income) %>%
#   distinct() %>%
#   select(-hidp)
# 
# # Step 2: Calculate income quintiles
# households_with_kids <- households_with_kids %>%
#   group_by(time) %>%
#   mutate(income_quintile = ntile(hh_income, 5)) %>%
#   ungroup()
# 
# # Step 3: Count households within each income quintile per year
# household_count_by_income_quintile <- households_with_kids %>%
#   group_by(time, income_quintile) %>%
#   summarise(n = n())
# 
# household_count_by_income_quintile
```

### Households With Children (Ethnic Group)

```{r}
raw.hhs_kids.ethnicity <- raw.dat %>%
  filter(nkids > 0) %>%
  select(hidp, time, hh_income, ethnicity) %>%  # Select relevant columns to reduce data size
  distinct() %>%
  group_by(time, ethnicity) %>%
  summarise(n = n(), .groups = 'drop') %>%
  mutate(source = 'final_US')

base.hhs_kids.ethnicity <- base.dat %>%
  filter(nkids > 0) %>%
  select(hidp, time, hh_income, ethnicity) %>%  # Select relevant columns to reduce data size
  distinct() %>%
  group_by(time, ethnicity) %>%
  summarise(n = n(), .groups = 'drop') %>%
  mutate(source = 'baseline')

combined <- rbind(raw.hhs_kids.ethnicity, base.hhs_kids.ethnicity)
combined$ethnicity <- factor(combined$ethnicity)

ggplot(combined, aes(x = time, y = n, group = ethnicity, color = ethnicity, fill = ethnicity)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = start.year) +
  labs(title = 'Households with Children Count')
```

```{r}
calculate_hh_proportions <- function(data, group.var) {
  # Step 1: Filter for households with children
  households_with_kids <- data %>%
    filter(nkids > 0) %>%
    select(hidp, time, hh_income, .data[[group.var]]) %>%
    distinct()

  # Step 3: Count households within each income group per year
  household_counts <- households_with_kids %>%
    group_by(time, .data[[group.var]]) %>%
    summarise(count = n(), .groups = 'drop')

  # Step 5: Calculate proportion of households in each group per year
  household_proportions <- household_counts %>%
    left_join(total_households_per_year, by = "time") %>%
    mutate(proportion = count / total_count) %>%
    select(time, .data[[group.var]], proportion)

raw.hh_kids_prop <- calculate_hh_proportions(raw.dat, group.var = 'ethnicity')
base.hh_kids_prop <- calculate_hh_proportions(base.dat, group.var = 'ethnicity')

combined <- rbind(raw.hh_kids_prop, base.hh_kids_prop)
combined$ethnicity <- factor(combined$ethnicity)

ggplot(combined, aes(x = time, y = proportion, group = ethnicity, color = ethnicity, fill = ethnicity)) +
  geom_col() +
  geom_vline(xintercept = start.year) +
  labs(title = 'Households with Children Count')
```
